---
title       : Predicting the amount of pollutants in any location within the US
subtitle    : Description of the shiny app "pred_pollutant_app"" 
author      : Oluwatobi Olabiyi
job         : Data Science
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

1. This presentation gives detailed desricption of the overall functionality and backend processing of the shining app "pred_pollutant_app"
2. The app can be found at https://engr3os.shinyapps.io/pred_pollutant_app
3. Distributed monitoring sensors are used to monitor polluntants amounts
4. Sometimes the pollution information of a particular location is not availble due to unavailability of monitoring sensor in that location
5. Pollution data from multiple sensors close to the location can be used used to predict the average amount of pollutant at the place of interest
 

--- .class #id 

## Pred_pollutant_app
The app was designed by taking the year 2014 pollutant data from EPA website http://aqsdr1.epa.gov/aqsweb/aqstmp/airdata. The data is subset and stored in data frame pollavg and it contains only longitude, latitude, pollutant name and the level of the pollutant. 
### Preprocessing
First few rows of data frame pollavg is given below:
```{r preproc, echo=FALSE, cache=TRUE}
url <- "http://aqsdr1.epa.gov/aqsweb/aqstmp/airdata/annual_all_2014.zip"
download.file(url, destfile = "annual_all_2014.zip")
unzip("annual_all_2014.zip")
d <- read.csv("annual_all_2014.csv")
## head(d)
sub = subset(d, Parameter.Name %in% c("PM2.5 - Local Conditions", "Ozone")
             & Pollutant.Standard %in% c("Ozone 8-Hour 2008", "PM25 Annual 2006"),
             c("Longitude", "Latitude", "Parameter.Name", "Arithmetic.Mean"))
pollavg <- aggregate(sub[, "Arithmetic.Mean"], 
                     sub[, c("Longitude", "Latitude", "Parameter.Name")], 
                     mean, na.rm = TRUE)
## recode the aggregated variable to name "Level"
names(pollavg)[4] <- "Level"
## reffacor the Parameter.Name variable to have only 2 levels
pollavg <- transform(pollavg, Parameter.Name = factor(Parameter.Name))
rm(d, sub)
monitors <- data.matrix(pollavg[, c("Longitude", "Latitude")])
save(pollavg, monitors, file = "polldata.RData")
head(pollavg)
```

---

## Average Pollutant Prediction of a Location

This section develops the algorithm to predict the average amount of pollutant of a location based on Roger Peng's yhat illustration in Data Product class Video.

This is the main section of the shinyapp pred_pollutant_app.

Given the logitude and latitude of a particular location, we can calculate the average PM2.5 and Ozone pollutants in an area of a particular radius.

```{r, echo=FALSE, comment = F}
library(fields)
library(xtable)
## Input is data frame with 
## lon: Longitude
## lat: Latitude
## radius: radius in miles

pollutant <- function(df){  
  x <- data.matrix(df[, c("lon", "lat")])
  r <- df$radius
  d <- rdist.earth(monitors, x)
  use <- lapply(seq_len(ncol(d)), function(i){
    which(d[, i] < r[i])
  })
  levels <- sapply(use, function(idx){
    with(pollavg[idx, ], tapply(Level, Parameter.Name, mean))
  })
  dlevels <- as.data.frame(t(levels))
  data.frame(df, dlevels)
}

```
For example, given the logitude and laitude of Baltimore, MD are -76.61 and 39.28 respectively. We can predict the average pollutant within 30 miles radious as:

```{r }
pollutant(data.frame(lon = -76.61, lat = 39.28, radius = 30))
```

---

## Further Details

Please note that the shiny app contains sliders that can be used to select longitude, latitude and radius parameters. 

The longitude and latitude have been limited to ranges with the continental US (-124.62 to -62.36 for longitude and 24.31 to 49.2 for latitude)
(Source: https://en.wikipedia.org/wiki/Extreme_points_of_the_United_States)

The radius is limited between 20 and 80 miles to provide reasonable prediction. 

For a particulatr location, with longitude and latitude, you can vary the radius to see the efect of including more or less monitors.

You can easily find the longitude and latitude of several US cities here: http://www.mapsofworld.com/lat_long/usa-lat-long.html